Approximate State Abstraction

### 2016-01-20 paper/paper.tex
Added intro and background to related work.

\documentclass{article}
\usepackage{times}
\usepackage{graphicx}
\usepackage{subfigure} 
\usepackage{natbib}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{subfigure}
\usepackage{hyperref}
\newcommand{\theHalgorithm}{\arabic{algorithm}} % Fixes misbehavior of conflicting packages.
\usepackage{icml2016/icml2016}

@@

@@

% --- ABSTRACT ---
\begin{abstract}
The combinatorial explosion that plagues planning and reinforcement-learning algorithms can be reversed using abstraction. For instance, prohibitively difficult task representations can be condensed so that solutions are tractably computable. In this work, we investigate a theoretical framework for approximate state abstraction that preserves near optimal behavior. Reinforcement learning agents using these abstractions may treat experiences that resemble each other as equivalent, and generalize knowledge to novel scenarios based on prior experiences. We present theoretical guarantees of the quality of value functions derived from four classes of approximate state abstraction. Additionally, we empirically evaluate the relationship between the degree of approximation and the degree of abstraction achieved, as well as the tradeoff between approximation magnitude and optimality of behavior.
\end{abstract}

@@

% --- SECTION: Introduction ---
\section{Introduction}

\begin{itemize}
\item Solving MDPs is P-Complete, where the dominant factor is $|\mathcal{S}|$.
\item Sample complexity is {\it also} dominated by $|\mathcal{S}|$.
\item To make MDPs with large $|\mathcal{S}|$ tractable, we can reduce the MDPs to a simpler form.
\item In this work, we show that compressing MDPs in a particular way allows for the transfer of bounded-error solutions between the compressed MDP and ground MDP.
\item Q: Why approximate?
\item A: Approximate can compress strictly more than compression based on equality
\item A: Approximate compression requires the type of knowledge that we could imagine learning (i.e.. approximate knowledge)
\end{itemize}

@@

\subsection{MDP/RL}

\subsection{Lihong Li}

@@

% Lihong/Walsh/Littman: Towards a unified theory...

@@

\citep{andre2002state} investigate the notion of {\it safe} state abstraction through a programming language called ALISP that can be used to specify hierarchical abstractions to reinforcement learning agents, enabling agents to ignore irrelevant parts of the state. The primary contribution of this work is the development of ALISP, and the decomposition of the Bellman equation into a form that enables safe planning algorithms using the abstractions from ALISP. Critically, our work differs in that our primary aim is to achieve understanding of what methods of abstraction in reinforcement learning can maintain bounded error solutions, while ALISP investigates one particular means of abstraction for use in planning, using partial programs.

@@

~\citep{Bean2011} develop an an algorithm (SADA) that performs aggregation on planning tasks that can be structured as a DAG, and conducts analysis on this process, achieving bounded error on the problem resulting from the aggregation. Our work differs in that the scope of aggregation and analysis is novel, and we make no assumptions about the structure of the underlying task. Lastly, we consider the reinforcement learning setting, as opposed to planning.

\item {\bf Approximate equivalence of MDP}~\cite{even2003approximate}

A number of works consider abstractions related to transition model similarity, or both transition model and reward function similarity with the aim of reducing the state space size required for planning and learning~\cite{even2003approximate,ortner2013adaptive,dean1997model,Boutilier2000}. This work is closely in line with our own in that we seek abstractions that may inform a reduced, abstract MDP, the solution to which has bounded error in the original MDP. However, we extend these results by considering broader classes of abstraction beyond the model similarity, or $\epsilon-homogeneity$, first introduced by ~\cite{dean1997model}.

\item {\bf Does approximate aggregation and then UCRL}~\cite{ortner2013adaptive}

Very similar to our overall goal: learn an abstraction function online using confidence bounds as the epsilon. Wowza. Only uses the whole model similarity business though. Seems to be the case for all of the approximate aggregation work (just does the bisimulation or $\epsilon-homogeneity$)

@@

@@

~\cite{jiang2015abstraction} investigate the question of choosing between two abstractions during learning. They provide analysis for an online approach to identifying the better abstraction.

>>>>

\documentclass{article}
\usepackage{times}
\usepackage{graphicx}
\usepackage{subfigure} 
\usepackage{natbib}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{subfigure}
\usepackage{hyperref}
\newcommand{\theHalgorithm}{\arabic{algorithm}} % Fixes misbehavior of conflicting packages.
\usepackage{icml2016/icml2016} 
\usepackage[nolist]{acronym} % Acronym package -- give argument <dua> to suppress acronym usage

@@

% --- Notation Commands ---
\newcommand{\ep}{\widetilde \phi}

@@

%Acronyms -- use \ac for acronym or \acp for plural/capitalization of acronym
\begin{acronym}
\acro{MDP}{Markov Decision Process}
\acrodefplural{MDP}[MDPs]{Markov Decision Processes}

\acro{RL}{reinforcement learning}
\acrodefplural{RL}{RL}{Reinforcement learning}
\end{acronym}

% --- ABSTRACT ---
\begin{abstract}
The combinatorial explosion that plagues planning and \ac{RL} algorithms can be reversed using abstraction. For instance, prohibitively difficult task representations can be condensed so that solutions are tractably computable. In this work, we investigate a theoretical framework for approximate state abstraction that preserves near optimal behavior. \acp{RL} agents using these abstractions may treat experiences that resemble each other as equivalent, and generalize knowledge to novel scenarios based on prior experiences. We present theoretical guarantees of the quality of value functions derived from four classes of approximate state abstraction. Additionally, we empirically evaluate the relationship between the degree of approximation and the degree of abstraction achieved, as well as the tradeoff between approximation magnitude and optimality of behavior.
\end{abstract}

@@

% --- SECTION: Introduction ---
\section{Introduction}
Abstraction is a fundamental tool of intelligent agent seeking to compute over only the salient features of its environment. However, the knowledge available to an agent at any given moment is typically approximate and so agents are faced with the challenge of deciding when their knowledge is sufficiently accurate to form the basis of abstraction. This work focuses on characterizing what ``sufficiently accurate" information for abstraction amounts to in the context of planning and \ac{RL} in \acp{MDP}.

Solving for optimal behavior in \acp{MDP} in a planning setting is known to be P-Complete in the size of the state space CITE MICHAEL COMPLEXITY PAPER. Similarly, many \ac{RL} algorithms for solving MDPs are known to require a number of samples polynomial in the size of the state space CITE SOME SAMPLE COMPLEXITY PAPER. Although polynomial runtime or sample complexity is often what one hopes, due to Bellman's so-called curse of dimensionality the state space of an MDP often grows super-polynomially with the number of variables that characterize the domain. Thus, polynomial behavior in state space size is not necessarily the \enote{bee's knees}. For instance, a robot involved in a pick and place task might be able employ planning algorithms to solve for how to manipulate some objects into a desired configuration in time polynomial in the number of states, but the number of states it must consider to do so grows exponentially with the number of objects it plans over CITE DABE ICAPS.

Thus, a key research interest of planning and RL has been leveraging abstraction as means of grappling with large state spaces (CITE VARIOUS ABSTRACTION JOANT). These methods are characterized by reducing \textit{ground} MDPs with large state spaces to \textit{abstract} MDPs with smaller state spaces by aggregating states with similar values of interest. Existing work has characterized how aggregation of states with equal values of particular quantities fully maintain optimality CITE LIHONG. However, exactly solving for these quantities is itself computationally taxing and sometimes as difficult as solving for optimal behavior in the ground MDP, thereby defeating the purpose of abstraction.

In this work we demonstrate that aggregation of states on the basis of various approximately equal criteria incur bounded error.  Relaxing the aggregation criteria from equality of quantities to epsilon-closeness offers two benefits. First, state aggregation algorithms that satisfy these criteria employ the sort of knowledge which we might reasonably a planning or learning algorithm to approximate without fully solving the MDP. Second, because the state aggregation criteria are relaxed to approximate equality, methods which employ approximate equality are able to tune the aggressiveness of state aggregation all the while incuring bounded error.

\enote{QUICK SUMMARY OF PAPER PARAGRAPH}

%\begin{itemize}
%\item Solving MDPs is P-Complete, where the dominant factor is $|\mathcal{S}|$.
%\item Sample complexity is {\it also} dominated by $|\mathcal{S}|$.
%\item To make MDPs with large $|\mathcal{S}|$ tractable, we can reduce the MDPs to a simpler form.
%\item In this work, we show that compressing MDPs in a particular way allows for the transfer of bounded-error solutions between the compressed MDP and ground MDP.
%\item Q: Why approximate?
%\item A: Approximate can compress strictly more than compression based on equality
%\item A: Approximate compression requires the type of knowledge that we could imagine learning (i.e.. approximate knowledge)
%\end{itemize}

@@

%MDP/SDM Background and Notation
\subsection{\acp{MDP} and Sequential Decision Making}
A \ac{MDP} is a five-tuple: $\langle \mathcal{S}, \mathcal{A}, \mathcal{T}, \mathcal{R}, \gamma \rangle$: $\mathcal{S}$ is a finite state space; $\mathcal{A}$ is the set of actions available to the agent; $\mathcal{T}$ denotes $\mathcal{T}(s' \mid s,a)$, the probability of an agent applying action $a \in \mathcal{A}$ in state $s \in \mathcal{S}$ and arriving in $s' \in \mathcal{S}$; $\mathcal{R}(s)$ denotes the reward received by the agent arriving in state $s$; $\gamma \in [0, 1]$ is a discount  factor that defines how much the agent prefers immediate rewards over future rewards.

The goal of an agent in an \ac{MDP} is to solve for a policy that maps states to actions: $\pi: \mathcal{S} \rightarrow \mathcal{A}$. In particular, the agent wants to solve for the policy that maximizes its expected discounted reward from any state. This goal policy is denoted $\pi^*$. We denote the expected discounted reward for following policy $\pi$ from state $s$ as the value of the state under that policy, $V(s)^\pi$. We similarly denote the expected discounted reward for taking action $a \in \mathcal{A}$ and then following policy $\pi$ from state $s$ forever after as $Q^\pi(s,a)$. Lastly, we denote the value and $Q$ functions under the optimal policy as $V^*$ and $Q^*$ respectively.

%Lihong Section
\subsection{Abstraction Notation}
We build upon the notation used by \cite{li2006towards} \enote{This should probably say more}.
We understand an abstraction as a mapping from a ground MDP $M_G$ to an abstract MDP $M_A$ using a state aggregation scheme. $M_G = \langle \mathcal{S}_G, \mathcal{A}, \mathcal{T}_G, \mathcal{R}_G, \gamma \rangle$ and $M_A = \langle \mathcal{S}_A, \mathcal{A}, \mathcal{T}_A, \mathcal{R}_A, \gamma \rangle$.

The states in the abstract MDP are constructed by applying a state aggregation scheme, $\phi$, to the states in the ground MDP. $\mathcal{S}_A$

@@

% Lihong/Walsh/Littman: Towards a unified theory...
Citation test~\cite{li2006towards}

@@

Programming language for enabling abstraction.

@@

Assumes the MDP is a DAG. Only does shortest path.

@@

\item DP with Factored Representations~\cite{Boutilier2000}

Uses decision trees and DBNs to represent MDPs. Does state aggregation. Theorem 4.4 should be looked at in more detail.

\item {\bf Model reduction techniques for nearly optimal MDP solutions}~\cite{dean1997model}

$\epsilon$-homogeneity: groups together states that behave approximately the same under all or some subset of policies.

Bounded Parameter MDP: family of MDP with upper and lower bounds on T. Find policies approximately optimal w.r.t. original MDP.
% Also covers model minimization..

@@
